---
title:  "Predictive modeling"
author: "Dennis Fok/Bas Donkers"
date:   "March, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load the data
```{r}
load("C:/Users/Palas/Documents/EUR/Text Analytics/reviews_with_predictor_variables.rData")
colnames(reviews_df) <- gsub(" ", "_", colnames(reviews_df))  # replace spaces in variable names
reviews_df <- reviews_df[, -96]     # drop NA column

reviews_df$rating <- as.numeric(reviews_df$rating)            # 0=happy, 1=not happy

N_factors <- 20
N_emotions <- 10 # includes pos/neg
N_words_stored <- 50
N_bigrams_stored <- 11
```

Get feature names and split sample
```{r Split sample}
index <- 8
factornames  <- colnames(reviews_df)[index:(index+N_factors-1)]
index <- index + N_factors
wordnames    <- colnames(reviews_df)[index:(index+N_words_stored-1)]
index <- index + N_words_stored
emotionnames <- colnames(reviews_df)[index:(index+N_emotions-1)]
index <- index + N_emotions
bigramnames <- colnames(reviews_df)[index:(index+N_bigrams_stored-1)]
index <- index + N_bigrams_stored

set.seed(1234)    # fix seed to allow for results to be reproducible
estimation_sample <- sort(sample(1:nrow(reviews_df), size = round(0.7*nrow(reviews_df))))
test_sample <- (1:nrow(reviews_df))[-estimation_sample]
```

Prepare some strings for use in models
```{r }
allFactors <- paste("(", paste(factornames,collapse=" + "), ")")
allEmotions <- paste("(", paste(emotionnames,collapse=" + "), ")")
allWords <- paste("(", paste(wordnames,collapse=" + "), ")")
allBigrams <- paste("(", paste(bigramnames,collapse=" + "), ")")
allWordsAndBigrams <- paste("(", paste(c(wordnames, bigramnames),collapse=" + "), ")")
# show example
allFactors
allEmotions
allWords
allBigrams
```

# Basic linear model (without interactions and variable selection)
```{r Linear Models}
f <- paste("rating ~ Nr_of_words + ", allFactors, " + ", allEmotions, " + ", allWords , " + ", allBigrams)
lm.all <- lm(f, data=reviews_df[estimation_sample,] )
summary(lm.all)

f <- paste("rating ~ Nr_of_words + ", allFactors, " + ", allWords , " + ", allBigrams)
lm.nodict <- lm(f, data=reviews_df[estimation_sample,] )
summary(lm.nodict)

f <- paste("rating ~  Nr_of_words + ", allFactors)
lm.onlyfactors <- lm(f, data = reviews_df[estimation_sample, ])
summary(lm.onlyfactors)

f <- paste("rating ~  Nr_of_words + ", allEmotions)
lm.onlyemotions <- lm(f, data = reviews_df[estimation_sample,])
summary(lm.onlyemotions)
 
f <- paste("rating ~ Nr_of_words + ", allWords)
lm.onlywords <- lm(f, data = reviews_df[estimation_sample,])
summary(lm.onlywords)

f <- paste("rating ~ Nr_of_words + ", allBigrams)
lm.bigrams <- lm(f, data = reviews_df[estimation_sample,])
summary(lm.bigrams)

f <- paste("rating ~ Nr_of_words + ", allWords , " + ",allBigrams)
lm.words_bigrams <- lm(f, data = reviews_df[estimation_sample,])
summary(lm.words_bigrams)

f <- paste("rating ~ Nr_of_words + positive + negative")
lm.posneg <- lm(f, data = reviews_df[estimation_sample,])
summary(lm.posneg)
```
```{r AIC comparison}
AIC(lm.all, lm.nodict, lm.onlyfactors, lm.onlyemotions, lm.onlywords, lm.bigrams, lm.words_bigrams, lm.posneg)
```
Best model is lowest AIC is model with all!


```{r test nested}
anova(lm.nodict, lm.all)
anova(lm.posneg, lm.onlyemotions)
anova(lm.words_bigrams, lm.all)
```

```{r Plot prediction histograms}
library(ggplot2)
dat <- data.frame(Prediction=predict(lm.all), rating=reviews_df[estimation_sample, "rating"])
ggplot(dat, aes(x=Prediction))  + 
    geom_histogram(data=subset(dat,rating == 0),fill = "red", alpha = 0.2) +
    geom_histogram(data=subset(dat,rating == 1),fill = "blue", alpha = 0.2)
```
Compare to logit model (GLM)
```{r glm}
f <- paste("(rating==1) ~ Nr_of_words + ", allFactors, " + ", allEmotions, " + ", allWords , " + ", allBigrams)
glm.all <- glm(f, data=reviews_df[estimation_sample,] , family=binomial)
summary(glm.all)
```


Plot GLM fit
```{r glm fit}
hist(predict(glm.all, type="response"))
dat <- data.frame(Predicted_prob=predict(glm.all, type="response"), rating=reviews_df[estimation_sample, "rating"])
ggplot(dat, aes(x=Predicted_prob))  + 
    geom_histogram(data=subset(dat,rating == 0),fill = "red", alpha = 0.2) +
    geom_histogram(data=subset(dat,rating == 1),fill = "blue", alpha = 0.2)

```


Compare fit: hitrate
```{r}
library(caret)
predglm <- as.factor(predict(glm.all, type="response") > .5)
predlm  <- as.factor(predict(lm.all, type="response") > .5)

confusionMatrix(data = predlm,  
                reference = as.factor(reviews_df[estimation_sample,"rating"]==1))

confusionMatrix(data = predglm,  
                reference = as.factor(reviews_df[estimation_sample,"rating"]==1))
```
```{r all word interactions}
f <- paste("rating ~  ", allWords,"^2 +", allBigrams)
lm.interactionswords <- lm(f, data = reviews_df[estimation_sample,])
summary(lm.interactionswords)
```

```{r interactions vs. bigrams}
AIC(lm.words_bigrams, lm.interactionswords)
anova(lm.interactionswords, lm.words_bigrams)
```

```{r Variable importance}
# on t-values
vi <- varImp(lm.all)
vi$Variable <- rownames(vi)
vi <- vi[order(-vi$Overall),]
vi$Variable <- factor(vi$Variable, levels = rev(vi$Variable))

ggplot(vi[1:25, ], aes(Variable,Overall)) + geom_bar(stat = "identity") + coord_flip() + ylab("Variable importance (t-value based)")

ggplot(vi[(nrow(vi)-24):nrow(vi), ], aes(Variable,Overall)) + geom_bar(stat = "identity") + coord_flip() + ylab("Variable importance (t-value based)")
```

```{r}
# on standardized coefficients
library("matrixStats")
vi <- coef(lm.all)/sqrt(colVars(model.matrix(formula(lm.all), data=reviews_df[estimation_sample,])))
StdCoef <- data.frame(StdCoef=vi[2:length(vi)])
StdCoef$Variable <- rownames(StdCoef)
StdCoef <- StdCoef[order(- StdCoef$StdCoef),]
StdCoef$Variable <- factor( StdCoef$Variable, levels = rev( StdCoef$Variable))

StdCoef <-  StdCoef[1:25,]
ggplot( StdCoef, aes(Variable,StdCoef)) + geom_bar(stat = "identity") + coord_flip() + ylab("Variable importance (standardized coefficients)")
```
```{r}
# on coef/ (max-min)
library("matrixStats")
X <- model.matrix(formula(lm.all), data=reviews_df[estimation_sample,])
vi <- coef(lm.all)/( apply(X, 2, max) - apply(X, 2, min) )
StdCoef <- data.frame(StdCoef=vi[2:length(vi)])
StdCoef$Variable <- rownames(StdCoef)
StdCoef <- StdCoef[order(- StdCoef$StdCoef),]
StdCoef$Variable <- factor( StdCoef$Variable, levels = rev( StdCoef$Variable))

StdCoef <-  StdCoef[1:25,]
ggplot( StdCoef, aes(Variable,StdCoef)) + geom_bar(stat = "identity") + coord_flip() + ylab("Variable importance (coefficients/range)")
```

Forward/backward selection is slow... do not do on large model
```{r}
f <- paste("rating ~  Nr_of_words + ", allFactors, " + ", allEmotions)
lm.fe_step <- step(lm(f, data = reviews_df[estimation_sample,]), direction = "both")
summary(lm.fe_step)
```

# Lasso regression
```{r Lasso}
library(glmnet)
library(plotmo) # for plot_glmnet
```
```{r}
f = paste("~ 0 + Nr_of_words + ", allFactors, " * ", allEmotions, " + ", allWords, " + ", allBigrams)

# Collect explanatory variables in a (large) matrix
LargeX <- model.matrix(formula(f), data=reviews_df)
y <- as.matrix(reviews_df[estimation_sample, "rating"])

lasso.mod <- glmnet(LargeX[estimation_sample,], y, alpha = 1)

plot_glmnet(lasso.mod) 
plot(lasso.mod)

cvfit <- cv.glmnet(LargeX[estimation_sample,], y, alpha = 1)
plot(cvfit)

coef(lasso.mod, cvfit$lambda.1se)

par <- predict(lasso.mod, s = cvfit$lambda.min, type='coefficients')
nnzero(par)
par <- predict(lasso.mod, s = cvfit$lambda.1se, type='coefficients')
nnzero(par)

lasso.pred <- predict(lasso.mod, s = cvfit$lambda.1se, newx = LargeX[estimation_sample,])
lasso.pred.test <- predict(lasso.mod, s = cvfit$lambda.1se, newx = LargeX[test_sample,])
mean((reviews_df[test_sample, "rating"]-lasso.pred.test)^2)
```

# Lasso logit = much slower...
```{r}
f = paste("~ 0 + Nr_of_words + ", allFactors, " * ", allEmotions, " + ", allWords, " + ", allBigrams)
LargeX <- model.matrix(formula(f), data=reviews_df)

y <- as.factor(reviews_df[estimation_sample, "rating"])
lasso.glm <- glmnet(LargeX[estimation_sample,], y, family="binomial", alpha = 1)

plot_glmnet(lasso.glm, nresponse = 5) 
plot(lasso.glm)

cvfit.glm <- cv.glmnet(LargeX[estimation_sample,], (y), family="binomial", alpha = 1)
plot(cvfit.glm)

coef(lasso.glm, cvfit.glm$lambda.1se)

par <- predict(lasso.glm, s = cvfit.glm$lambda.min, type='coefficients')
nnzero(par)
length(par)

lasso.glm.pred <- predict(lasso.glm, s = cvfit.glm$lambda.1se, type="response", newx = LargeX[estimation_sample,])
lasso.glm.pred.test <- predict(lasso.glm, s = cvfit.glm$lambda.1se,type="response", newx = LargeX[test_sample,])
```

```{r Linear with Interactions}
f <- paste("rating ~  Nr_of_words + ", allFactors,"+", allEmotions,"+", allWordsAndBigrams, " + ", allFactors, " *", allEmotions, " + ", allFactors , "* ", allWordsAndBigrams, " + ", allEmotions , " * ", allWordsAndBigrams)
lm.interactions <- lm(f, data = reviews_df[estimation_sample,])
summary(lm.interactions)
```


```{r}
AIC(lm.all, lm.onlyfactors, lm.onlyemotions, lm.onlywords, lm.interactions, lm.fe_step)
```

Predictive performance
```{r}
for (m in list(lm.all, lm.onlyfactors, lm.onlyemotions, lm.onlywords, lm.interactions, lm.fe_step))
{
  predicted <- predict.lm(m, reviews_df[estimation_sample,])
  mse <- mean((as.numeric(reviews_df[estimation_sample,"rating"])-predicted)^2)
  
  predicted_test <- predict.lm(m, reviews_df[test_sample,])
  mse_test<- mean((as.numeric(reviews_df[test_sample,"rating"])-predicted_test)^2)
  print(c(sqrt(mse), sqrt(mse_test)))
}
print("lm.all, lm.onlyfactors, lm.onlyemotions, lm.onlywords, lm.interactions, lm.fe_step")
```

# Random forest
```{r}
library("randomForest")
```
```{r}
f = paste("as.factor(rating) ~ Nr_of_words + ", allFactors, " * ", allEmotions, " + ", allWords, " + ", allBigrams)
rf = randomForest(formula(f),  
                   ntree = 100,
                   data = reviews_df[estimation_sample,])
plot(rf)
print(rf)
```

```{r}
varImpPlot(rf,  
           sort = T,
           n.var = 20,
           main = "Top 20 - Variable Importance")
```

```{r}
library(caret)
pred.alt <- predict(rf)
print("estimation_oob")
confusionMatrix(data = pred.alt,  
                reference = as.factor(reviews_df[estimation_sample,"rating"]))


pred.est <- predict(rf, reviews_df[estimation_sample,])
print("estimation_insample")
confusionMatrix(data = pred.est,  
                reference = as.factor(reviews_df[estimation_sample,"rating"]))

pred.test <- predict(rf, reviews_df[test_sample,])
print("validation")
confusionMatrix(data = pred.test,  
                reference = as.factor(reviews_df[test_sample,"rating"]))
```

```{r forecast comparison}
for (m in list(lm.all, lm.nodict, lm.onlyfactors, lm.onlyemotions, lm.onlywords, lm.bigrams, lm.words_bigrams, lm.posneg, lm.interactions, lm.fe_step, lm.interactionswords))
{
  predicted <- 1+(predict.lm(m, reviews_df[estimation_sample,]) >= 0.5)
  predicted_test <- 1+ (predict.lm(m, reviews_df[test_sample,]) >= 0.5)
  hit <- mean(predicted == reviews_df[estimation_sample,"rating"])
  hit.test <- mean(predicted_test == reviews_df[test_sample,"rating"])
  print(c(hit,hit.test))
}
print("lm.all, lm.nodict, lm.onlyfactors, lm.onlyemotions, lm.onlywords, lm.bigrams, lm.words_bigrams, lm.posneg, lm.interactions, lm.fe_step, lm.interactionswords")

print("GLM")
pred.glm <- predict(glm.all, type="response") >= 0.5
pred.glm.test <- predict(glm.all, type="response", newdata=reviews_df[test_sample,]) >= 0.5
print("estimation_insample")
mean(pred.glm == (reviews_df[estimation_sample,"rating"]==1))
print("test")
mean(pred.glm.test == (reviews_df[test_sample,"rating"]==1))

print("Lasso LM")
pred <- 1+(lasso.pred >= 0.5)
hit <- mean(pred == reviews_df[estimation_sample,"rating"])
pred_test <- 1+(lasso.pred.test >= 0.5)
hit.test <- mean(pred_test == reviews_df[test_sample,"rating"])
print(c(hit,hit.test))

print("Lasso GLM")
hit <- mean((lasso.glm.pred>=0.5) == (reviews_df[estimation_sample,"rating"]==1))
hit.test <- mean((lasso.glm.pred.test>=0.5) == (reviews_df[test_sample,"rating"]==1))
print(c(hit,hit.test))


print("RF")
pred.alt <- predict(rf)
print("estimation_oob")
mean(pred.alt == as.factor(reviews_df[estimation_sample,"rating"]))


pred.est <- predict(rf, reviews_df[estimation_sample,])
print("estimation_insample")
mean(pred.est == as.factor(reviews_df[estimation_sample,"rating"]))

pred.test <- predict(rf, reviews_df[test_sample,])
print("test")
mean(pred.test == as.factor(reviews_df[test_sample,"rating"]))
```
# END